---
title: "4. Fantastic Bursts and how to use them"
output:
  pdf_document: default
  html_document: default
vignette: |
  %\VignetteIndexEntry{3. Working with an sftrack}
   %\VignetteEncoding{UTF-8}
   %\VignetteEngine{knitr::rmarkdown}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::load_all("/home/matt/r_programs/sftrack")
#library(sftrack)

```

## Bursts  

Bursts are a big emphasis in the class. They are made in a similar vein to the `sfc` and `sfg` in `sf`. `ind_burst` is a singular burst. Its whats stored at the row level. 
A `multi-burst` is a collection of `ind_bursts` and exists at the column level. Bursts also have an `active_burst` argument, which turns on and off certain bursts for analysis and plotting purposes.

We can look at the structure
```{r}
burst_list = list(id = data$sensor_code, month = as.POSIXlt(data$acquisition_time)$mon+1)

mb1 <- make_multi_burst(x=burst_list, active_burst=c('id','month'))
str(mb1)
mb1[[1]]
```
A burst contains grouping information, where the id of the subject is required, but additional bursts are not. Every burst has the same burst columns available, and these bursts are stored internally as a list. You can also see that labels are created based on the active_burst, this can be accessed and makes for easy labeling of plots of figures.  

#### Basics
##### ind_bursts 
An ind_burst is the grouping variables for a single row of data.  

You can make an `ind_burst` object using `make_ind_burst()`, and giving it a list with the bursts named. 

```{r}

indb <- make_ind_burst(list(id='CJ13', month = 4), active_burst = c('id','month'))
str(indb)
```

Because ind_bursts are simply lists, you can edit individual elements in an ind_burst
```{r}
indb 
indb[1] <- 'CJ15'
indb$month <- '5'
str(indb)
```

ind_bursts is a grouping class where you can have different groups 'turned on' as the analysis allows. Which groups are activated is determined by the 'active burst'. You can set this when creating an ind_burst, otherwise it defaults to all groups being activated at once.

```{r}
indb <- make_ind_burst(list(id='CJ13', month = 4), active_burst = c('id'))
str(indb)
indb <- make_ind_burst(list(id='CJ13', month = 4))
str(indb)

````

You can access or change the `active burst` at anytime using the `active_burst()` function. The active burst tells any analysis to group by that mix of variables. 

```{r}
active_burst(indb)
active_burst(indb) <- 'id'
str(indb)

```

To speed up this process by calculating a 'label' attribute is calculated for each ind_burst which is a combination of the active burst variables. This label is to increase calculation and subsetting speed, and is therefore not available for renaming. It can be accessed however by `burst_labels`. This can be useful when refering to a burst or selecting many bursts in a column

```{r}
burst_labels(indb)
burst_labels(mb1)[1:10]
```
##### multi_burst  
Multi_bursts are a collection of ind_bursts where all ind_bursts have the same active_burst.

Similarly to ind_burst you can make a multi_burst with `make_multi_burst()`. The argument `burst` takes a list where each element is a vector indicating the named burst as well as a vector of the active bursts.

```{r}
burst_list <- list(id = rep(1:2,10), year = rep(2020, 10))
mb <- make_multi_burst(x=burst_list, active_burst=c('id','year'))
str(mb)
```

You can also make a multi_burst by concatenating multiple ind_bursts. Though in this case the active_burst will default to all bursts. Which you can change later.

```{r}
a <- make_ind_burst(list(id = 1, year = 2020))
b <- make_ind_burst(list(id = 1, year = 2021))
c <- make_ind_burst(list(id = 2, year = 2020))
mb <- c(a, b , c)

summary(mb)
```

You can also combine multi_bursts together with `c()`.  

```{r}
mb_combine <- c(mb,mb)
summary(mb_combine)
```
You can also edit bursts like a list, but you must replace it with an object of the appropriate class
```{r}
mb[1]
mb[1] <- make_ind_burst(list(id=3,year=2019))
mb[1]

```

And the burst names must match the ones in the multi_burst
```{r}
# Try to add an ind_burst with a month field when the original burst had year instead 
tryCatch(mb[1] <- make_ind_burst(list(id=3,month=2019)), error = function(e) e)
```

##### selecting a burst

As multibursts are stored as lists, it can be difficult to refer to a single burst or group of bursts. This is where the burst label can come in handy. The burst label is remade everytime an active burst changes, and therefore can be used to subset. Using `burst_label()` you can access the variable to subset with. Additionally the label comes with a factor argument, that returns a factor instead of a character. But it should be noted the variable is stored internally as a character in each ind_burst. So the factor returned may not remain the same if a burst has changed.

```{r}
burst_list <- list(id = rep(1:2,10), year = rep(2020, 10))
mb <- make_multi_burst(x=burst_list, active_burst=c('id','year'))
burst_labels(mb)[1:10]

head(my_sftraj[burst_labels(my_sftraj) %in% c('CJ13'), ])


```

You can also subset by entering the burst label in the burst itself in either the multi_burst or the sftrack/sftraj object:

```{r}
mb['1_2020']

sub <- my_sftrack['CJ13_1',]
print(sub,5, 3)
```
##### active_burst 
The active_burst is a simple yet powerful feature. It dictates how your data is grouped for essentially all calculations. It can also be changed on the fly. You can also view and change the active_burst of a multi_burst with `active_burst()`. This calculates a new sort_index internally as well.

```{r}
active_burst(my_sftrack$burst)
summary(my_sftrack, stats = T)
active_burst(my_sftrack$burst) <- c('id')
active_burst(my_sftrack$burst)
summary(my_sftrack, stats = T)

```
