---
title: "sftraj working vignette"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
devtools::load_all("/home/matt/r_programs/sftraj")

```

Load data and create appropriate variables for import. We'll set 0's to NA manually for now.  

**data** is a data frame  
**burst** is a list with multiple named vectors to group the data by  
**active_burst** is a vector of the burst names you want to be activated for analysis  
**error** is a vector of the error data  
**time** is a vector of POSIXct or integer for the time  
**coords** is a vector of the columns for the x,y,z coordinates  

```{r}
data("raccoon_data")
raccoon_data$latitude[raccoon_data$latitude==0] <- NA
raccoon_data$longitude[raccoon_data$longitude==0] <- NA
raccoon_data$height[raccoon_data$height==0] <- NA
burst = list(id = raccoon_data$sensor_code,month = as.POSIXlt(raccoon_data$utc_date)$mon, height =as.numeric(raccoon_data$height>5))
active_burst = c('id','month')
error = rep(NA, nrow(raccoon_data))
time = as.POSIXct(raccoon_data$acquisition_time, tz = 'UTC')
coords = c('latitude','longitude','height')
```

###############################################

Starting with the simple stuff. `as_sftrack` is the generic function to create an object. Can accept vectors, sf + vectors, and ltraj  
```{r}
# sftrack
my_track <- as_sftrack(data = raccoon_data, burst = burst, active_burst = active_burst, error = error, time = time , coords = coords )
my_track
# sfstep
my_step <- as_sfstep(data = raccoon_data, burst = burst, active_burst = active_burst, error = error, time = time , coords = coords )
my_step
```

This generic function also converts ltraj objects  
```{r, message = FALSE}
library(adehabitatLT)

ltraj_df <- as.ltraj(xy=raccoon_data[,c('longitude','latitude')], date = as.POSIXct(raccoon_data$acquisition_time),
 id = raccoon_data$sensor_code, typeII = TRUE,
 infolocs = raccoon_data[,1:6] )

as_sftrack(ltraj_df)
as_sfstep(ltraj_df)
```


### inter class conversion
This function also converts between the `sftrack` and `sfstep` data type

```{r}
as_sftrack(my_step)
as_sfstep(data=my_track)
```


## Plotting

Currently there are some plotting methods. Base plotting is basic, but assumes the active_burst is the grouping/coloring variable  

```{r}
plot(my_step)
```


You can also plot in ggplot, but right now its only using the geom_sftraj function. We can figure this out at a later date. But the method is very simple,
you just have to tell geom_sf() what the grouping variable is.

```{r}
library(ggplot2)
ggplot() + geom_sftraj(data = my_step)
```

## Bursts  

Bursts are a big point in the class. They are made in a similar vain to the `sfc` and `sfg` in `sf`. `ind_burst` is a singular birst. Its whats
stored at the row level. A `multi-burst` is a collection of `ind_bursts` and exists at the column level. Bursts also have an `active_burst` argument, which turns on and off certain bursts for analysis and plotting purposes.


```{r}
mb1 <- make_multi_burst(burst=burst, active_burst=c('id','month'))
str(mb1)
```


Burst select allows you access to the active_burst

```{r}
burst_select(mb1)[1]
```

burst labels allows access to the burst labels which are a combination of the active bursts with '_' seperator

```{r}
burst_labels(mb1)[1:10]
```

 
Can combine multiple ind_bursts to create a multi_burst
```{r}
mb2 <- c(ind_burst(list(id = 1, year = 2018)),ind_burst(list(id = 2, year = 2018))) 
str(mb2)
```
**The rest is not done quite complete** 
And can redefine burst levels
