---
title: "5. Getting Spatial with sftrack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Getting Spatial with sftrack}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    fig.width = 6,
    fig.asp = 0.7
)
library("lwgeom")

```

Let's first check the two `sftrack` and `sftraj` objects provided in
the `raccoons` dataset:

```{r}
library("sftrack")
racc_track
racc_traj

```


## Geometry column

As stated earlier, the geometry column is built using `sf`, so the column functions exactly as it would in `sf`. You can modify it and redefine it using the `sf` tools. More specifically the geometry column of an `sftrack` object is an `sfc` column. The main difference between a standard `sf` object created using `st_as_sf` is that we automatically allow empty geometries, where as this option is turned off by default in `st_as_sf()`.

```{r}
racc_track$geometry

```

An `sftrack` object is simply an `sfc` of `sfc_POINTS`, this contrasts
with an `sftraj` object which is a mixture of `POINT`s and
`LINESTRING`s. This is because a trajectory can have a start point and
a missing end point (NA), a line segment, or a missing start point and
an actual end point. As linestrings can not have NULL points in them,
this allows no-loss conversion back and forth between `sftrack` and an
`sftraj`.

```{r}
racc_traj$geometry

```

This does mean that not all `sf` functions will handle an `sftraj` object like it would an `sftrack` if there are NAs in the data set. In these cases `st_is_empty()` can help to subset the points that contain geometry data.


## Working with sf

As an `sftrack` object is an `sf` object, essentially all of the `sf` functions will work on it.

```{r}
library("sf")
st_length(racc_traj)[1:10]

df1 <- data.frame(
  id = c(1,1,1,1),
  time = as.POSIXct('2020-01-01 12:00:00', tz = 'UTC') + 60*60*(1:4),
  x = c(1,3,3,2),
  y = c(1,1,3,4)
)

road <- st_linestring(rbind(
  c(1,2),
  c(5,2),
  c(5,0)
)
)

animal1 <- as_sftraj(df1)
plot(animal1)
plot(road,  col = 'red', add = T)

# Does the animal cross the road?
any(st_intersects(animal1, road, sparse = FALSE))

# When?
animal1$time[st_intersects(animal1, road, sparse = FALSE)]

# How often does the animal stay near the road?
st_is_within_distance(animal1, road, 1)

# How close is the animal from the road?
st_distance(animal1, road)

```

The only thing to remember, is that a `sftraj` has a `GEOMETRY`
column, and occasionally a function may not work with it. In those
cases `is_linestring()` can be used to filter out incomplete steps
(either missing, or without start or end point).


## Plotting


### Base plotting 

`sftrack` is built to work with `sf` base plot methods. This means you can use most of the `sf` plot methods, `sftrack` largely just controls the grouping of the plot then feeds it back to `plot.sf()`. 

```{r, fig.show = "hold", out.width = "50%"}
plot(racc_traj)
plot(racc_traj)

```

This means that everytime you change the active_group, the plot view will change.

```{r}
active_group(racc_traj$sft_group) <- "id"
active_group(racc_traj$sft_group)
plot(racc_traj)

```

Most arguments for `plot.sf` are available to use as additional arguments to `plot`. 

```{r}
plot(racc_track, axes = TRUE, cex = 5)

```


### ggplot2

This is a work in progress, but theres a `geom_sftrack()` function
that feeds `geom_sf()` with the correct plotting information. Like
with `geom_sf()`, you input `data` into `geom_sftrack()` and not into
`ggplot()`.  Here too, `ggplot2` assumes `active_group` is the grouping variable. Plots vary slightly based on if they're `sftrack` of `sftraj` class.

`geom_sftrack()` is essentially `geom_sf(data = data, aes(color = group_labels(data)))` with NULL points subetted out. This may help when a user requires more advanced modification than the `geom_sftrack()` allows.

```{r}
library("ggplot2")
ggplot() + geom_sftrack(data = racc_traj)

```

You can use `geom_sftrack()` just like any other `ggplot2` layer, which means you can continue to make manuscript quality plots.

```{r}
cols <- c("TTP-041_1" = "dodgerblue1", "TTP-041_2" = "darkseagreen2",
          "TTP-058_1" = "darkorchid1", "TTP-058_2" = "khaki3")
ggplot() +
  geom_sftrack(data = racc_track, size = 3, alpha = 0.5) +
  scale_color_manual(values = cols) + 
  ggtitle("Raccoons at Tree Tops Park, Winter 2020") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Special functions for working with an sftraj

To help with working with more complex sftraj objects, there is a growing suite of `sftraj` specific functions:


### Step mode in plotting

`sftraj` objects create linestrings for each row of data, as each row is not assumed to be related to the next row of data. This data structure may become inefficient to work with when plotting large data sets. When appropriate these trajectories can be generalized by combining linestrings where lines meet.

For plotting purposes we can create these linestrings quickly and plot them at much faster speeds than individual lines. 

For `plot` and `geom_sftrack` there is an argument called `step_mode`
that refers to whether you want to plot the lines individually
(`step_mode = TRUE`), or generalize them into connected linestrings
(`step_mode = FALSE`). By default `step_mode` is set to `TRUE` for
small datasets (less than 10,000 records), and switches to `FALSE` for
larger datasets to enhance performance.

```{r, fig.show = "hold", out.width = "50%"}
library("ggplot2")
ggplot() + geom_sftrack(data = racc_traj, step_mode = TRUE) +
  ggtitle("Step mode enabled")
ggplot() + geom_sftrack(data = racc_traj, step_mode = FALSE) +
  ggtitle("Step mode disabled")

```

You'll notice that the appearance of the plot changes as the POINTs are also displayed. This is because step mode adds POINTS to the plot that contain a fill as well as a color property.


### `coord_traj`

This function returns a `data.frame` (x,y,z) of the point at $t_1$ of each `sftraj` geometry. It works nearly identically to `sf::st_coordinates()`.

```{r}
coord_traj(racc_traj$geometry)[1:10, ]

```


### `pts_traj `

If youd like to retain the geometries but still pull out $t_1$ point you can use `pts_traj()`. This functions returns a list of the beginning point of each `sftraj` geometry, or an `sfc` column when using the argument `sfc = TRUE`.

```{r}
pts_traj(racc_traj$geometry)[1:5]
pts_traj(racc_traj$geometry, sfc = TRUE)[1:5]

```

### `is_linestring`

This function may help if you'd like to quickly filter an `sftraj` object to just contain pure linestrings. `is_linestring()` returns TRUE or FALSE if the geometry is a linestring. This does not recalculate anything, it just filters out steps that contained NAs in either phase. Its nearly identical to `st_is(x, "LINESTRING")`, but may be more intuitive for users.

```{r}
is_linestring(racc_traj$geometry)[1:10]
(new_sftraj <- racc_traj[is_linestring(racc_traj$geometry), ])

```


### Calculating step metrics of an `sftraj`

For use in movement models, you may need to calculate the $dx$, $dy$, length, and turn angles of each step. You can do that in `sftrack` using `step_metrics()`. It should be noted it will accept an `sftrack` object, however, it first converts the geometries internally to `sftraj` geometries and then calculates step metrics. As with other `sf` objects, the outcome is assumed to be in the units of the `crs` when not specified. Absolute angle is measured in radians.

```{r}
step_metrics(racc_traj[1:10,])

```
